### 2024-03-27 paper/paper.qmd
updated results and llm usage

---
title: "Evaluating the Influence of Education on Current Unemployment Trends"
author: 
  - Adrian Ly
thanks: "Code and data are available at: [LINK](https://github.com/AdrianUofT/unemployment-employment-rate)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(readr)
library(lubridate)
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(arrow)
library(rstanarm)
library(modelsummary)
library(lme4)

unemployed_data <- read.csv(here::here("data/analysis_data/cleaned_unemployed_data.csv"))
```

# Introduction

# Data

```{r}
#| echo: false
#| eval: true
#| label: tbl-data
#| tbl-cap: "Unemployed Dataset from 2019 to 2023 by education level"
#| warning: false
#| message: false
#| results: 'asis'

# Customizing column names
column_names <- c("Education Level", "Reference Period", "Youth Unemployment", "Adult Unemployment", "Senior Unemployment")

# Assuming your data is in a dataframe called unemployed_data
# Display the head of the dataframe with custom column names
kable(head(unemployed_data), col.names = column_names, caption = "Unemployed Dataset from 2019 to 2023 by Education Level")
```

## Data Measurement

## Data Analysis

This paper uses the Statistics of Canada `Unemployment rate, participation rate and employment rate by educational attainment, annual` [@citedataset] dataset. We are using the `R programming language` [@citeR] to conduct our analysis along with `readR` [@citereadR], `lubridate` [@citelubridate], `tidyverse` [@citetidyverse], `dplyr` [@citedplyr], `tidyr` [@citetidyr], `knitr` [@citeknitr], `janitor` [@citejanitor], `scales` [@citescales], `RColorBrewer` [@citeRColorBrewer], `ggplot2` [@citeggplot2], `kableExtra` [@citekableExtra], `here` [@citeHere], `arrow` [@citeArrow], `rstanarm` [@citerstanarm], `modelsummary` [@citemodelsummary], and `lme4` [@citelme4].

# Model

The goal of our modeling strategy, utilizing multiple linear regression, is multifaceted:

- **Understanding Relationships**: To understand the relationship between several independent variables, such as education level and age groups, and the dependent variable, which in this context is the youth unemployment rates. The model is designed to isolate the effect of each variable on unemployment rates.

- **Quantifying Impact**: To quantify the impact of each predictor. Regression coefficients provide a numerical value that represents the expected change in the dependent variable for a one-unit change in an independent variable, all else being equal.

- **Predictive Analysis**: To develop a predictive framework that can be employed to estimate or forecast the outcome variable, in this case, unemployment rates, when provided with new data for the predictors.

- **Inferential Statistics**: To conduct hypothesis testing to ascertain whether the relationships observed within the data are statistically significant and not merely due to random variation.

- **Policy Decision Support**: To offer empirical evidence that may inform policy decisions. For instance, if a correlation is found between higher education levels and increased unemployment, it might indicate the necessity for policies that better align education with employment opportunities.

- **Data-Driven Insights**: To extract insights from the data that could lead to an enhanced understanding of labor market dynamics, such as how different age groups' unemployment rates interact.

- **Modeling Assumptions Testing**: To validate the assumptions underlying the regression model, ensuring the reliability and accuracy of the model's predictions.

- **Confounding Variable Control**: To account for potential confounders that might influence the relationship between education level and unemployment rates. The inclusion of relevant variables aims to minimize estimation bias.

- **Sensitivity Analysis**: To comprehend the model's sensitivity to changes in inputs, which assists in evaluating the robustness of the model.

This strategy is designed to produce a model that is both explanatory and predictive, offering a current snapshot of the relationships in the data as well as a tool for future forecasting. Such a model can be utilized to guide decision-making processes, or pinpoint where interventions may be effective.

## Model set-up

$$
\begin{aligned}
\text{Youth Unemployment} = \beta_0 &+ \beta_1 \cdot \text{Education Level} \\
&+ \beta_2 \cdot \text{Reference Period} \\
&+ \beta_3 \cdot \text{Adult Unemployment} \\
&+ \beta_4 \cdot \text{Senior Unemployment} + \epsilon
\end{aligned}
$$

Where:

- `Youth Unemployment` is the dependent variable we are trying to predict. These are generally categorized as people who are between the ages of 15-24.
- `Education Level` is an independent variable representing the level of education. (e.g., high school, undergraduate, postgraduate) 
- `Reference Period` is another independent variable that represents the time frame of the data collection (e.g., year).
- `Adult Unemployment` and `Senior Unemployment` are independent variables representing unemployment rates for adult and senior age groups, with people ages ranging between 25-44 and above 45, respectively.
- $\beta_0$ is the y-intercept, representing the expected value of `Youth Unemployment` when all the independent variables are 0.
- $\beta_1, \beta_2, \beta_3, \beta_4$ are the coefficients for each independent variable, representing the change in `Youth Unemployment` for a one-unit change in the respective independent variable, holding all other variables constant.
- $\epsilon$ represents the error term, accounting for the variability in `Youth Unemployment` not explained by the model.

This model aims to understand how various factors contribute to youth unemployment. The coefficients 
$\beta_1, \beta_2, \beta_3, \beta_4$ provide insights into the relationship between each independent variable and youth unemployment:

- $\beta_1$ tells us how changes in the education level might affect youth unemployment, holding other factors constant. A positive coefficient suggests that higher education levels are associated with higher youth unemployment rates, while a negative coefficient suggests the opposite.

- $\beta_2$ gives insight into how changes over the reference period (e.g., years) impact youth unemployment, which can help identify trends over time.

- $\beta_3$ and $\beta_4$ help understand how unemployment rates in other age groups (adults and seniors) are related to youth unemployment. This can reveal whether higher unemployment rates in these groups are associated with higher or lower youth unemployment rates.

The intercept $\beta_0$ provides the baseline level of youth unemployment when all other variables are zero, which may not always be a practical scenario but is necessary for the model's mathematical formulation. The error term $\epsilon$ captures the model's limitations and any random factors affecting youth unemployment that aren't included in the model.

## Model justification

The utilization of a multiple linear regression model for analyzing the relationship between people's unemployment status, age groups, and education level over time is substantiated by various rationales. Primarily, the model adeptly accommodates multiple independent variables such as education level and age group, each potentially exerting influence on the dependent variable, unemployment status. This capacity allows for an assessment of the individual impact of each predictor while simultaneously controlling for the influence of others. Such a model not only facilitates the quantification of the relationships between these factors and unemployment—via coefficients that express the extent of change in response per unit change in a predictor (with other variables held constant)—but also enables hypothesis testing on the parameters. This testing is instrumental in determining whether the associations between predictors and the response are statistically significant, thereby offering insights into the meaningfulness of these relationships.

# Results

## Model Interpretation:

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

model <-
  readRDS(file = here::here("models/unemployed_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Regression Analysis: Impact of Education Level and Age on Youth Unemployment"
#| warning: false
#| message: false
#| results: 'asis'

# Specify custom names for coefficients and goodness-of-fit statistics if needed
custom_coef_names <- c(
  "education_levelAbove bachelor's degree" = "Above Bachelor's",
  "education_levelBachelor's degree" = "Bachelor's",
  "education_levelHigh school graduate" = "High School Grad",
  "education_levelPostsecondary certificate or diploma" = "Postsec. Cert/Dip",
  "education_levelSome high school" = "Some High School",
  "education_levelSome postsecondary" = "Some Postsec.",
  "education_levelTotal, all education levels" = "All Levels",
  "education_levelUniversity degree" = "University",
  "reference_period" = "Reference Period",
  "adult_unemployment" = "Adult Unemployment",
  "senior_unemployment" = "Senior Unemployment"
)

# Use modelsummary to create the table
modelsummary::modelsummary(
  list(
    "Baseline Model for Youth Unemployment" = model
  ),
  coef_rename = custom_coef_names,
  stars = TRUE,
  fmt = 2,
)
```

Based on the regression model output, as seen in @tbl-modelresults, several interpretations can be drawn:

- **Baseline Unemployment**: The intercept of `r format(328.01, nsmall = 2)` suggests a baseline youth unemployment level, conditional on all other predictor variables being zero.

- **Impact of Higher Education**:
  - *Above Bachelor’s Degree*: The coefficient `r format(5.14, nsmall = 2)`*** indicates a significant and positive association with youth unemployment rates, implying potential overqualification or job market mismatches.
  - *Bachelor’s Degree*: Similarly, a coefficient of `r format(4.14, nsmall = 2)`** suggests that a bachelor's degree is associated with higher unemployment.
  - *University Degree*: With a coefficient of `r format(4.27, nsmall = 2)`**, university graduates also face higher unemployment rates.

- **Some Postsecondary Education**: A negative coefficient of `r format(-2.20, nsmall = 2)`+ for those with some postsecondary education indicates a decrease in unemployment compared to the reference group.

- **Adult Unemployment**: The significant positive coefficient `r format(2.19, nsmall = 2)`*** for adult unemployment implies a correlation with youth unemployment rates.

- **Time Factor**: The non-significant coefficient for the reference period suggests that time does not significantly alter youth unemployment rates when holding other factors constant.

- **Goodness of Fit**:
  - The high R-squared value of `r format(0.923, nsmall = 3)` indicates that the model explains much of the variance in youth unemployment rates.
  - The Adjusted R-squared value of `r format(0.898, nsmall = 3)` takes into account the number of predictors and suggests a good fit to the data.

- **Significance Levels**:
  - The asterisks (`***`, `**`, `*`) represent the level of statistical significance, with more stars denoting stronger evidence against the null hypothesis.

These findings underscore the complexity of youth unemployment and the varied impact of educational attainment on job market outcomes.

## Youth Unemployment

The results presented in @tbl-youth2019 and @tbl-youth2022 shows the percentage change of youth unemployment rates over the years by education level for 2019-2023 and 2022-2023 respectively. The column on the far right shows the net percentage change for that time period.The following formula was used to calculate the net percentage change in the unemployment rate:

$$
\text{Net Percentage Change} = \left( \frac{\text{Value at End Period} - \text{Value at Start Period}}{\text{Value at Start Period}} \right) \times 100\%
$$

```{r}
#| echo: false
#| eval: true
#| label: tbl-youth2019
#| tbl-cap: "Percentage Change in Youth Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(youth_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-youth2022
#| tbl-cap: "Percentage Change in Youth Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(youth_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempyouth
#| fig-cap: Youth Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = youth_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Youth Unemployment by Education Level Over Time",
       x = "Year",
       y = "Youth Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

Looking at @fig-unempyouth, it gives us a visualization of the youth unemployment rates for varying education levels across the years. We then used the following formula to calculate the overall percentage change between years:

$$
\text{Overall Percentage Change} = \left( \frac{\text{Sum of Rates in X Year} - \text{Sum of Rates in Y Year}}{\text{Sum of Rates in Y Year}} \right) \times 100\%
$$

The overall percentage change for the youth unemployment rate is $\approx 0.92 \%$ increase between 2019 and 2023 and $\approx 10.45 \%$ increase between 2022 and 2023.

The figures and tables from our analysis present a nuanced picture of youth unemployment across different educational strata over a period extending from 2019 to 2023.

- **Trends Over Time**: A clear pattern of fluctuation is evident in the unemployment rates. The peak observed in 2020 may correlate with the economic downturn caused by the COVID-19 pandemic, affecting youth employment disproportionately.

- **Impact of Education**: The variation in unemployment by education level underscores the importance of educational attainment. Notably, higher levels of education such as "Above Bachelor’s" and "Bachelor’s" do not necessarily correlate with lower unemployment rates, suggesting a potential oversupply of higher education qualifications or mismatches between education and job market needs.

- **Percentage Changes**:
  - *Net Percentage Change*: Individual educational levels show diverse trends, with some experiencing increases in unemployment, while others show decreases over the four-year span.
  - *Overall Percentage Change*: The aggregate data reveals a modest uptick in youth unemployment rates, which may reflect broader economic and societal shifts.

## Adult Unemployment

```{r}
#| echo: false
#| eval: true
#| label: tbl-adult2019
#| tbl-cap: "Percentage Change in Adult Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(adult_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-adult2022
#| tbl-cap: "Percentage Change in Adult Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(adult_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempadult
#| fig-cap: Adult Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = adult_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Adult Unemployment by Education Level Over Time",
       x = "Year",
       y = "Adult Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

Similar to the youth unemployment, @tbl-adult2019 and @tbl-adult2022 show the net percentage change in the adult unemployment for the years 2019-2023 and 2022-2023 respectively.

The analysis of adult unemployment rates from 2019 to 2023 uncovers distinct trends across education levels. Notably, individuals with "0 to 8 years" of education show a marked increase in unemployment rates, which could indicate that lower educational attainment is a significant risk factor for adult unemployment.

Conversely, adults with education "Above bachelor's degree" experienced a decrease in unemployment rates, suggesting that higher education may offer better protection against job market downturns. This trend aligns with the expectation that higher qualifications translate to greater employment stability.

For other education categories, such as "Bachelor's degree" and "High school graduate," the changes are relatively moderate, indicating a more nuanced relationship between education and job market outcomes for these groups.

Overall, the adult unemployment rate shows an approximately $\approx 0.89 \%$ increase from 2019 to 2023, which might reflect the lingering effects of economic events or shifts in the labor market dynamics. However, the period from 2022 to 2023 sees a more pronounced increase of approximately $\approx 9.98 \%$, highlighting the potential for rapid changes in unemployment within a single year.

These patterns underscore the importance of ongoing education and training throughout one's career, as well as the need for targeted support for adults with lower levels of education.

@fig-unempadult illustrates the fluctuation in unemployment rates across different education levels. It reveals that unemployment is not uniform across the educational spectrum, which has implications for policymakers and educators alike.

## Senior Unemployment

```{r}
#| echo: false
#| eval: true
#| label: tbl-senior2019
#| tbl-cap: "Percentage Change in Senior Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(senior_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-senior2022
#| tbl-cap: "Percentage Change in Senior Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(senior_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempsenior
#| fig-cap: Senior Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = senior_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Senior Unemployment by Education Level Over Time",
       x = "Year",
       y = "Senior Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

The study of senior unemployment sheds light on the intricate interplay between education levels and job market viability. Key observations include:

- **Divergent Patterns**: A pronounced decrease in unemployment rates was observed among seniors with vocational training as seen from @tbl-senior2019 and @tbl-senior2022, while those with higher academic degrees saw an uptick, raising questions about the alignment between higher education and job opportunities in the market.

- **Consistency Among University Graduates**: Seniors holding university degrees experienced a marginal increase in unemployment, suggesting a relative consistency within this group compared to others.

- **Overall Shifts**: There was an overall reduction in senior unemployment from 2019 to 2023 with a decrease of approximately  $\approx 13.54 \%$ and a decrease of approximately $\approx 6.11 \%$ from 2022 to 2023, pointing towards potential shifts in retirement age, economic recoveries, or evolving job opportunities for seniors.

- The overarching trends in senior unemployment call for strategic educational planning and job placement efforts tailored to the nuanced needs of the aging workforce.

- The plotted data shown in @fig-unempsenior provides a visual narrative of unemployment trends among seniors, highlighting that educational achievements translate differently into job security for the senior population.

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# References