### 2024-03-29 paper/paper.qmd
updated paper and references

---
title: "Evaluating the Influence of Education on Current Unemployment Trends"
author: 
  - Adrian Ly
thanks: "Code and data are available at: [LINK](https://github.com/AdrianUofT/unemployment-employment-rate)"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(readr)
library(lubridate)
library(tidyverse)
library(dplyr)
library(tidyr)
library(knitr)
library(janitor)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(kableExtra)
library(here)
library(arrow)
library(rstanarm)
library(modelsummary)
library(lme4)

unemployed_data <- read.csv(here::here("data/analysis_data/cleaned_unemployed_data.csv"))
```

# Introduction

# Data

```{r}
#| echo: false
#| eval: true
#| label: tbl-data
#| tbl-cap: "Unemployed Dataset from 2019 to 2023 by education level"
#| warning: false
#| message: false
#| results: 'asis'

# Customizing column names
column_names <- c("Education Level", "Reference Period", "Youth Unemployment", "Adult Unemployment", "Senior Unemployment")

# Assuming your data is in a dataframe called unemployed_data
# Display the head of the dataframe with custom column names
kable(head(unemployed_data), col.names = column_names, caption = "Unemployed Dataset from 2019 to 2023 by Education Level")
```

The dataset outlined in @tbl-data captures a comprehensive view of unemployment rates spanning from 2019 to 2023, segmented by different education levels. This rich dataset provides insights into the unemployment trends for three distinct age groups: youth, adults, and seniors, highlighting how economic fluctuations impact various demographics over time. The granularity of the data allows for an in-depth analysis of how education, as a key socio-economic factor, correlates with employment stability and job market resilience during periods of economic change.

## Data Measurement

Our data set is comprised of data from the `Labour Force Survey` [@citedatasetsource]. The Labour Force Survey (LFS) is a principal source for the measurement of Canada's labour market activities. Here, we provide an overview of the survey methodology and the robustness of the data collected.

- Survey Design and Methodology

  - The LFS uses a stratified multi-stage sampling approach to gather data from a representative sample of the non-institutionalized civilian population aged 15 and over. The survey design meticulously excludes certain populations, such as full-time military personnel and residents of institutions, ensuring the accuracy and relevance of the labour force estimates.

- Collection and Processing

  - Conducted monthly, the LFS is characterized by its promptness, with data releases occurring just 10 days post-collection. The survey yields the early indicator estimates of employment and unemployment, alongside other vital labour market indicators including the employment rate and participation rate.

- Utilization of Data

  - Governments and policymakers extensively use LFS data to evaluate and plan employment programs. The data are essential for determining eligibility and the level of Employment Insurance benefits, as well as for the analysis conducted by economists, labour market analysts, and researchers.

- Supplementary Surveys

  - In conjunction with the LFS, surveys like the SEPH, EIS, and JVWS contribute to a comprehensive understanding of current labour market conditions, providing a more detailed and nuanced picture.

- Data Collection Techniques

  - Interviews are conducted through computer-assisted methods, either by telephone or in-person visits. This allows for the collection of a diverse set of labour market data, which is further refined through editing and imputation to account for non-responses or inconsistencies.

- Estimation Procedures

  - Sample data are meticulously weighted to produce reliable estimates at the national, provincial, and sub-provincial levels. These weights are adjusted to compensate for non-responses and to align the survey estimates with known population controls.

- Quality Control

  - To ensure the highest data quality, the LFS is subject to rigorous validation against other economic data sources. This includes the comparison of employment estimates with those from SEPH, EIS, and census data.

- Confidentiality and Disclosure Control

  - Adhering to stringent confidentiality rules, the LFS ensures that no individual or business information is disclosed without consent. The survey employs a range of suppression rules to maintain the anonymity of the data.

- Seasonal Adjustment

  - The LFS implements the X-12-ARIMA technique for seasonal adjustment, allowing for the analysis of short-term labour market trends without the influence of seasonal patterns.

- Accuracy and Reliability

  - Despite being subject to sampling errors inherent in survey data, the LFS employs coefficients of variation and other statistical measures to assess and enhance the reliability of its estimates.

This survey's methodology and the rapid release of its results make the LFS a cornerstone in the analysis and understanding of Canada's labour market dynamics.

## Data Analysis

This paper uses the Statistics of Canada `Unemployment rate, participation rate and employment rate by educational attainment, annual` [@citedataset] dataset. We are using the `R programming language` [@citeR] to conduct our analysis along with `readR` [@citereadR], `lubridate` [@citelubridate], `tidyverse` [@citetidyverse], `dplyr` [@citedplyr], `tidyr` [@citetidyr], `knitr` [@citeknitr], `janitor` [@citejanitor], `scales` [@citescales], `RColorBrewer` [@citeRColorBrewer], `ggplot2` [@citeggplot2], `kableExtra` [@citekableExtra], `here` [@citeHere], `arrow` [@citeArrow], `rstanarm` [@citerstanarm], `modelsummary` [@citemodelsummary], and `lme4` [@citelme4].

# Model

The goal of our modeling strategy, utilizing multiple linear regression, is multifaceted:

- **Understanding Relationships**: To understand the relationship between several independent variables, such as education level and age groups, and the dependent variable, which in this context is the youth unemployment rates. The model is designed to isolate the effect of each variable on unemployment rates.

- **Quantifying Impact**: To quantify the impact of each predictor. Regression coefficients provide a numerical value that represents the expected change in the dependent variable for a one-unit change in an independent variable, all else being equal.

- **Predictive Analysis**: To develop a predictive framework that can be employed to estimate or forecast the outcome variable, in this case, unemployment rates, when provided with new data for the predictors.

- **Inferential Statistics**: To conduct hypothesis testing to ascertain whether the relationships observed within the data are statistically significant and not merely due to random variation.

- **Policy Decision Support**: To offer empirical evidence that may inform policy decisions. For instance, if a correlation is found between higher education levels and increased unemployment, it might indicate the necessity for policies that better align education with employment opportunities.

- **Data-Driven Insights**: To extract insights from the data that could lead to an enhanced understanding of labor market dynamics, such as how different age groups' unemployment rates interact.

- **Modeling Assumptions Testing**: To validate the assumptions underlying the regression model, ensuring the reliability and accuracy of the model's predictions.

- **Confounding Variable Control**: To account for potential confounders that might influence the relationship between education level and unemployment rates. The inclusion of relevant variables aims to minimize estimation bias.

- **Sensitivity Analysis**: To comprehend the model's sensitivity to changes in inputs, which assists in evaluating the robustness of the model.

This strategy is designed to produce a model that is both explanatory and predictive, offering a current snapshot of the relationships in the data as well as a tool for future forecasting. Such a model can be utilized to guide decision-making processes, or pinpoint where interventions may be effective.

## Model set-up

$$
\begin{aligned}
\text{Youth Unemployment} = \beta_0 &+ \beta_1 \cdot \text{Education Level} \\
&+ \beta_2 \cdot \text{Reference Period} \\
&+ \beta_3 \cdot \text{Adult Unemployment} \\
&+ \beta_4 \cdot \text{Senior Unemployment} + \epsilon
\end{aligned}
$$

Where:

- `Youth Unemployment` is the dependent variable we are trying to predict. These are generally categorized as people who are between the ages of 15-24.
- `Education Level` is an independent variable representing the level of education. (e.g., high school, undergraduate, postgraduate) 
- `Reference Period` is another independent variable that represents the time frame of the data collection (e.g., year).
- `Adult Unemployment` and `Senior Unemployment` are independent variables representing unemployment rates for adult and senior age groups, with people ages ranging between 25-44 and above 45, respectively.
- $\beta_0$ is the y-intercept, representing the expected value of `Youth Unemployment` when all the independent variables are 0.
- $\beta_1, \beta_2, \beta_3, \beta_4$ are the coefficients for each independent variable, representing the change in `Youth Unemployment` for a one-unit change in the respective independent variable, holding all other variables constant.
- $\epsilon$ represents the error term, accounting for the variability in `Youth Unemployment` not explained by the model.

This model aims to understand how various factors contribute to youth unemployment. The coefficients 
$\beta_1, \beta_2, \beta_3, \beta_4$ provide insights into the relationship between each independent variable and youth unemployment:

- $\beta_1$ tells us how changes in the education level might affect youth unemployment, holding other factors constant. A positive coefficient suggests that higher education levels are associated with higher youth unemployment rates, while a negative coefficient suggests the opposite.

- $\beta_2$ gives insight into how changes over the reference period (e.g., years) impact youth unemployment, which can help identify trends over time.

- $\beta_3$ and $\beta_4$ help understand how unemployment rates in other age groups (adults and seniors) are related to youth unemployment. This can reveal whether higher unemployment rates in these groups are associated with higher or lower youth unemployment rates.

The intercept $\beta_0$ provides the baseline level of youth unemployment when all other variables are zero, which may not always be a practical scenario but is necessary for the model's mathematical formulation. The error term $\epsilon$ captures the model's limitations and any random factors affecting youth unemployment that aren't included in the model.

## Model justification

The utilization of a multiple linear regression model for analyzing the relationship between people's unemployment status, age groups, and education level over time is substantiated by various rationales. Primarily, the model adeptly accommodates multiple independent variables such as education level and age group, each potentially exerting influence on the dependent variable, unemployment status. This capacity allows for an assessment of the individual impact of each predictor while simultaneously controlling for the influence of others. Such a model not only facilitates the quantification of the relationships between these factors and unemployment—via coefficients that express the extent of change in response per unit change in a predictor (with other variables held constant)—but also enables hypothesis testing on the parameters. This testing is instrumental in determining whether the associations between predictors and the response are statistically significant, thereby offering insights into the meaningfulness of these relationships.

# Results

## Model Interpretation:

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

model <-
  readRDS(file = here::here("models/unemployed_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Regression Analysis: Impact of Education Level and Age on Youth Unemployment"
#| warning: false
#| message: false
#| results: 'asis'

# Specify custom names for coefficients and goodness-of-fit statistics if needed
custom_coef_names <- c(
  "education_levelAbove bachelor's degree" = "Above Bachelor's",
  "education_levelBachelor's degree" = "Bachelor's",
  "education_levelHigh school graduate" = "High School Grad",
  "education_levelPostsecondary certificate or diploma" = "Postsec. Cert/Dip",
  "education_levelSome high school" = "Some High School",
  "education_levelSome postsecondary" = "Some Postsec.",
  "education_levelTotal, all education levels" = "All Levels",
  "education_levelUniversity degree" = "University",
  "reference_period" = "Reference Period",
  "adult_unemployment" = "Adult Unemployment",
  "senior_unemployment" = "Senior Unemployment"
)

# Use modelsummary to create the table
modelsummary::modelsummary(
  list(
    "Baseline Model for Youth Unemployment" = model
  ),
  coef_rename = custom_coef_names,
  stars = FALSE,
  fmt = 2,
)
```

Based on the regression model output, as seen in @tbl-modelresults, several interpretations can be drawn:

- **Baseline Unemployment**: The intercept of `r format(328.01, nsmall = 2)` suggests a baseline youth unemployment level, conditional on all other predictor variables being zero.

- **Impact of Higher Education**:
  - *Above Bachelor’s Degree*: The coefficient `r format(5.14, nsmall = 2)`*** indicates a significant and positive association with youth unemployment rates, implying potential overqualification or job market mismatches.
  - *Bachelor’s Degree*: Similarly, a coefficient of `r format(4.14, nsmall = 2)`** suggests that a bachelor's degree is associated with higher unemployment.
  - *University Degree*: With a coefficient of `r format(4.27, nsmall = 2)`**, university graduates also face higher unemployment rates.

- **Some Postsecondary Education**: A negative coefficient of `r format(-2.20, nsmall = 2)`+ for those with some postsecondary education indicates a decrease in unemployment compared to the reference group.

- **Adult Unemployment**: The significant positive coefficient `r format(2.19, nsmall = 2)`*** for adult unemployment implies a correlation with youth unemployment rates.

- **Time Factor**: The non-significant coefficient for the reference period suggests that time does not significantly alter youth unemployment rates when holding other factors constant.

- **Goodness of Fit**:
  - The high R-squared value of `r format(0.923, nsmall = 3)` indicates that the model explains much of the variance in youth unemployment rates.
  - The Adjusted R-squared value of `r format(0.898, nsmall = 3)` takes into account the number of predictors and suggests a good fit to the data.

- **Significance Levels**:
  - The asterisks (`***`, `**`, `*`) represent the level of statistical significance, with more stars denoting stronger evidence against the null hypothesis.

These findings underscore the complexity of youth unemployment and the varied impact of educational attainment on job market outcomes.

## Youth Unemployment

The results presented in @tbl-youth2019 and @tbl-youth2022 shows the percentage change of youth unemployment rates over the years by education level for 2019-2023 and 2022-2023 respectively. The column on the far right shows the net percentage change for that time period.The following formula was used to calculate the net percentage change in the unemployment rate:

$$
\text{Net Percentage Change} = \left( \frac{\text{Value at End Period} - \text{Value at Start Period}}{\text{Value at Start Period}} \right) \times 100\%
$$

```{r}
#| echo: false
#| eval: true
#| label: tbl-youth2019
#| tbl-cap: "Percentage Change in Youth Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(youth_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-youth2022
#| tbl-cap: "Percentage Change in Youth Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(youth_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempyouth
#| fig-cap: Youth Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = youth_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Youth Unemployment by Education Level Over Time",
       x = "Year",
       y = "Youth Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

Looking at @fig-unempyouth, it gives us a visualization of the youth unemployment rates for varying education levels across the years. We then used the following formula to calculate the overall percentage change between years:

$$
\text{Overall Percentage Change} = \left( \frac{\text{Sum of Rates in X Year} - \text{Sum of Rates in Y Year}}{\text{Sum of Rates in Y Year}} \right) \times 100\%
$$

The overall percentage change for the youth unemployment rate is $\approx 0.92 \%$ increase between 2019 and 2023 and $\approx 10.45 \%$ increase between 2022 and 2023.

The figures and tables from our analysis present a nuanced picture of youth unemployment across different educational strata over a period extending from 2019 to 2023.

- **Trends Over Time**: A clear pattern of fluctuation is evident in the unemployment rates. The peak observed in 2020 may correlate with the economic downturn caused by the COVID-19 pandemic, affecting youth employment disproportionately.

- **Impact of Education**: The variation in unemployment by education level underscores the importance of educational attainment. Notably, higher levels of education such as "Above Bachelor’s" and "Bachelor’s" do not necessarily correlate with lower unemployment rates, suggesting a potential oversupply of higher education qualifications or mismatches between education and job market needs.

- **Percentage Changes**:
  - *Net Percentage Change*: Individual educational levels show diverse trends, with some experiencing increases in unemployment, while others show decreases over the four-year span.
  - *Overall Percentage Change*: The aggregate data reveals a modest uptick in youth unemployment rates, which may reflect broader economic and societal shifts.

## Adult Unemployment

```{r}
#| echo: false
#| eval: true
#| label: tbl-adult2019
#| tbl-cap: "Percentage Change in Adult Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(adult_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-adult2022
#| tbl-cap: "Percentage Change in Adult Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(adult_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempadult
#| fig-cap: Adult Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = adult_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Adult Unemployment by Education Level Over Time",
       x = "Year",
       y = "Adult Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

Similar to the youth unemployment, @tbl-adult2019 and @tbl-adult2022 show the net percentage change in the adult unemployment for the years 2019-2023 and 2022-2023 respectively.

The analysis of adult unemployment rates from 2019 to 2023 uncovers distinct trends across education levels. Notably, individuals with "0 to 8 years" of education show a marked increase in unemployment rates, which could indicate that lower educational attainment is a significant risk factor for adult unemployment.

Conversely, adults with education "Above bachelor's degree" experienced a decrease in unemployment rates, suggesting that higher education may offer better protection against job market downturns. This trend aligns with the expectation that higher qualifications translate to greater employment stability.

For other education categories, such as "Bachelor's degree" and "High school graduate," the changes are relatively moderate, indicating a more nuanced relationship between education and job market outcomes for these groups.

Overall, the adult unemployment rate shows an approximately $\approx 0.89 \%$ increase from 2019 to 2023, which might reflect the lingering effects of economic events or shifts in the labor market dynamics. However, the period from 2022 to 2023 sees a more pronounced increase of approximately $\approx 9.98 \%$, highlighting the potential for rapid changes in unemployment within a single year.

These patterns underscore the importance of ongoing education and training throughout one's career, as well as the need for targeted support for adults with lower levels of education.

@fig-unempadult illustrates the fluctuation in unemployment rates across different education levels. It reveals that unemployment is not uniform across the educational spectrum, which has implications for policymakers and educators alike.

## Senior Unemployment

```{r}
#| echo: false
#| eval: true
#| label: tbl-senior2019
#| tbl-cap: "Percentage Change in Senior Unemployment by education level (2019-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2019, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(senior_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2019`)/`2019` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2019)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-senior2022
#| tbl-cap: "Percentage Change in Senior Unemployment by education level (2022-2023)"
#| warning: false

# Filter data for the years 2019 and 2023
filtered_data <- unemployed_data %>%
  filter(reference_period %in% c(2022, 2023))

# Compare youth unemployment rate by education level for 2019 and 2023
comparison <- filtered_data %>%
  group_by(education_level, reference_period) %>%
  summarise(mean_unemployment = mean(senior_unemployment, na.rm = TRUE)) %>%
  pivot_wider(names_from = reference_period, values_from = mean_unemployment) %>%
  mutate(Percentage_Change = round((`2023` - `2022`)/`2022` * 100, 2)) %>%
  arrange(education_level)

kable(comparison, 
      digits = 2, 
      col.names = c("Education Level", "Unemployment Rate (2022)", "Unemployment Rate (2023)", "Percentage Change (%)"))
```

```{r}
#| label: fig-unempsenior
#| fig-cap: Senior Unemployment Rates
#| echo: false
#| warning: false
#| message: false

ggplot(unemployed_data, aes(x = reference_period, y = senior_unemployment, group = education_level, color = education_level)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Senior Unemployment by Education Level Over Time",
       x = "Year",
       y = "Senior Unemployment Rate (%)",
       color = "Education Level") +  # This labels the legend title
  scale_color_brewer(palette = "Set1") +
  scale_color_discrete(labels = c("0 to 8 years" = "Up to 8 Years",
                                  "Above bachelor's degree" = "Postgraduate",
                                  "Bachelor's degree" = "Bachelor's",
                                  "High school graduate" = "High School",
                                  "Postsecondary certificate or diploma" = "Postsec. Cert./Dip.",
                                  "Some high school" = "Partial High School",
                                  "Some postsecondary" = "Some College",
                                  "Total, all education levels" = "All Levels",
                                  "University degree" = "University"))

```

The study of senior unemployment sheds light on the intricate interplay between education levels and job market viability. Key observations include:

- **Divergent Patterns**: A pronounced decrease in unemployment rates was observed among seniors with vocational training as seen from @tbl-senior2019 and @tbl-senior2022, while those with higher academic degrees saw an uptick, raising questions about the alignment between higher education and job opportunities in the market.

- **Consistency Among University Graduates**: Seniors holding university degrees experienced a marginal increase in unemployment, suggesting a relative consistency within this group compared to others.

- **Overall Shifts**: There was an overall reduction in senior unemployment from 2019 to 2023 with a decrease of approximately  $\approx 13.54 \%$ and a decrease of approximately $\approx 6.11 \%$ from 2022 to 2023, pointing towards potential shifts in retirement age, economic recoveries, or evolving job opportunities for seniors.

- The overarching trends in senior unemployment call for strategic educational planning and job placement efforts tailored to the nuanced needs of the aging workforce.

- The plotted data shown in @fig-unempsenior provides a visual narrative of unemployment trends among seniors, highlighting that educational achievements translate differently into job security for the senior population.

# Discussion

## What we learn from these results

### The Impact of Education on Employment

Our data distinctly reveals that education significantly influences employability. The variation in unemployment rates across educational levels underlines the fact that not all education is equal in the eyes of the labor market. Higher education levels often correlate with lower unemployment rates, suggesting that advanced skills and qualifications can provide an edge in securing employment. This aligns with the longstanding view that educational attainment can enhance an individual's job prospects, potentially leading to more stable and higher-quality employment opportunities.

However, the data also highlight that higher education does not guarantee immunity against unemployment. This could be due to a myriad of factors, including the mismatch between the skills acquired through higher education and the needs of the job market, or possibly the saturation of certain degrees in the labor force. A great example of this is the computer science job market. People trying to get into the computer science job market are faced with the harsh reality of not being able to find a job. In fact, "Computer Science Students Face a Shrinking Big Tech Job Market" (@citesource1) according to the NY Times with "Hiring in tech, information and media is at its lowest level since July 2020" (@citesource2). With these graduate students seeking to find a job, they are most likely to go into an "entry-level position", if they even exist regardless of the field they are pursuing. The emphasis here being that these entry-level positions are either not existent, or the people working these jobs are majorly over-qualified. This is not even accounting for the number of layoffs that happen on a regular basis from these major tech companies. A very recent major layoff is from Bell Communications where they layed off "hundreds of employees in 10-minute video call meetings" (@citesource3) and that "no union reps or members were allowed to unmute themselves during the call and ask anything" (@citesource4). Thus, while education is a powerful tool for improving employment prospects, it is not a panacea. There is a need for ongoing evaluation of educational curricula to ensure alignment with evolving industry demands.

### The Resilience of Different Demographics

The study also sheds light on the resilience of various demographic groups in the face of economic fluctuations. Individuals with vocational training, in particular, demonstrate robustness in maintaining employment, potentially due to the practical nature of their skills and the specific demand for trades in the economy. Similarly, those with higher education—despite facing some challenges—often show a lesser impact from economic downturns, indicating that higher education may still provide a buffer against the vagaries of the job market.

These observations suggest that the job market values not just education but the relevance and applicability of the skills obtained. In other words, people are granted "senorities", hence why they are seen as more valuable to a company. As economies evolve and new industries emerge, the adaptability of the workforce becomes crucial. Hence, continuous learning and skills development are imperative for both individuals and economies to thrive.

## Why are we seeing these results

The patterns observed in unemployment rates across educational levels may be attributable to several underlying factors. Understanding why we see these results requires delving into the interconnections between educational systems, economic structures, and labor market demands.

### Educational System Alignment with Market Needs

A potential reason for the variation in unemployment rates could be the degree to which education systems are aligned with current market needs. Educational institutions that closely work with industries and update curricula to reflect technological advancements and market trends may better equip students with marketable skills. On the other hand, a mismatch between graduates' skills and job requirements can lead to higher unemployment rates, even among those with advanced degrees.

### Economic Shifts and Job Market Evolution

The global economy is in a constant state of flux, influenced by technological change, globalization, and, more recently, by the impact of the COVID-19 pandemic. These shifts can create or eliminate entire industries, affecting employment opportunities differently across educational levels. For instance, the growing emphasis on technology might benefit those with relevant qualifications while displacing jobs in sectors less adaptable to these changes. Which leads into my next point.

### The Influence of Artificial Intelligence on the Job Market

The impact of artificial intelligence (AI) on the job market is multifaceted and profound, acting as both a disruptor and a catalyst within the labor landscape. AI’s capacity for automating tasks has resulted in the displacement of certain jobs, particularly those involving routine and repetitive functions. However, it’s equally notable for creating new job categories that necessitate human oversight and management of these advanced systems. As AI redefines roles, the urgency for upskilling becomes apparent—workers are finding themselves in need of new competencies that align with the evolving demands of an AI-integrated workplace. The skills that are becoming increasingly indispensable are those that AI cannot easily replicate, such as complex problem-solving, critical thinking, and creative capacities. These skills highlight the enduring value of human ingenuity in the age of automation.

Furthermore, AI's contribution to economic efficiency and productivity has the potential to spur economic growth, indirectly fostering job creation with a shift in required skill-sets. The onus is on policymakers to navigate this transition, balancing the embrace of technological innovation with the need for robust support systems for the workforce. They face the challenge of creating policies that not only protect workers from the immediate risks of displacement but also encourage the development of industries poised to grow in the AI era. Ultimately, the integration of AI into the economy presents a complex narrative: while it may render some occupations obsolete, it simultaneously opens doors to new realms of employment. The key challenge—and opportunity—lies in harnessing AI as a tool to enhance human work, not replace it, by anticipating changes and preparing the workforce through education and training to thrive alongside intelligent machines.

### Job outsourcing

The globalization of the labor market has seen an increasing trend in companies outsourcing work internationally, driven by the quest for cost reduction, efficiency gains, and access to a broader talent pool. Outsourcing has become a strategic move for businesses looking to streamline operations and focus on core competencies while delegating peripheral activities to external specialists. This shift has had a significant impact on the job market.

Outsourced jobs often transition from higher-cost labor markets to regions where labor costs are lower, fundamentally altering the employment landscape in both the originating and recipient countries. While outsourcing can lead to job losses in some sectors within developed economies, it simultaneously generates employment opportunities in developing regions. This redistribution of jobs raises questions about the long-term sustainability of such practices, considering the potential for economic disparity and the need for international labor standards.

The implications for the domestic workforce are considerable. Jobs that are more vulnerable to outsourcing, particularly those that can be performed remotely or do not require extensive localized knowledge, have seen a decline. Conversely, jobs that necessitate a physical presence or rely heavily on local market understanding tend to remain domestically anchored. This dynamic has put a spotlight on the need for workers in developed economies to adapt to the changing market, often requiring a shift toward more specialized, technical, or management-focused roles that are less susceptible to outsourcing.

Companies are not only outsourcing for economic reasons but also to capitalize on global expertise. As certain regions develop specializations in fields like IT, customer service, or manufacturing, companies may outsource to benefit from high-quality services while also reducing costs.

### The Role of Vocational Training

Vocational training often provides specific skills directly tied to existing jobs in the economy. This training can lead to immediate employment upon completion, reflecting the lower unemployment rates observed for individuals with vocational qualifications. This suggests that vocational training may be more responsive to immediate labor market demands compared to traditional academic paths.

### Policy and Economic Support Structures

Government policies and economic support structures also play a critical role in employment outcomes. Economic stimulus measures, job creation programs, and educational subsidies can all influence employment rates. Regions with proactive workforce development policies might show more robust employment figures, particularly for demographics that benefit most from such policies.

## What could this entail for the future

Looking ahead, the tension between the advantages of outsourcing and its impact on the job market will persist as a key issue. It is imperative for policy to address the repercussions for workers while supporting economic growth and maintaining competitive advantages in a globalized economy. This includes investing in education and vocational training to equip the domestic workforce with skills that are in demand and less likely to be outsourced. Additionally, there is a growing call for multinational corporations to consider the ethical implications of their outsourcing strategies and strive for practices that are equitable and sustainable across all regions.

## Weaknesses and next steps

Weaknesses and next steps should also be included.

# Conclusion

\newpage

# References